---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = TRUE
)
library(gitpins)
stopifnot(!file.exists(here::here("gitpins")))
```

# gitpins

<!-- badges: start -->
[![R-CMD-check](https://github.com/torfason/gitpins/workflows/R-CMD-check/badge.svg)](https://github.com/torfason/gitpins/actions)
<!-- badges: end -->

`r desc::desc_get("Description")`

## The problem

You want to quickly and easily process an online resource using R functions, some of which only accept local files. Thus you would like the following properties for your workflow:

* Download to a local file
* But avoid downloading on every single run
* Refresh your data regularly from the online source
* Use a local copy if the online resource is not accessible
* Have the local copy be easily accessible in a predictable location
* Not ruin your local copy if the online version should change in a "bad" way

## The solution

The `gitpins` package downloads a URL to a local file in the `gitpins` folder
inside your project (the currently fixed path is determined by
`here("gitpins")`), and then returns the full file name name of the local file,
which can be passed as an argument to any function that expects to read such a
file.

```{r}
# Downloads on first try
gitpin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv") |> 
  read.csv() |> head()
```

You can maintain as many resources as you need:

```{r}
# Another resource
gitpin("https://vincentarelbundock.github.io/Rdatasets/csv/datasets/sunspot.month.csv") |> 
  read.csv() |> head()
```

The file is downloaded the first time you run `gitpin` on a given URL (the
actual download is done with `curl::curl_download()`). After that, it checks to
see the age of the local file and re-downloads if it is to old. The default
refresh interval is 12 hours, but is configurable with a parameter.

Note that the return value of the `gitpin()` function is simply the full path to
the local copy of the file. You can therefore use `gitpin()` with the original
URL wherever you would have used the local path of the resource. The exact name
of the file is constructed in a deterministic way based on the URL
(specifically, the base name is the `digest()` of the URL).

```{r}
# Uses a cached copy if a recent one is available (start of the url changed for privacy)
gitpin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv") |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")
```

The refresh interval is configured with the `refresh_hours` parameter. Use
`refresh_hours=0` to force a download on every call, and `refresh_hours=Inf` to
always use the local copy (after the first download).

```{r}
# Force a reload by specifying zero refresh time
gitpin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv",
       refresh_hours = 0) |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")

# Always use local copy by specifying Inf refresh time
gitpin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv",
       refresh_hours = Inf) |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")
```

The `gitpins` directory is actually a local `git` repository, and each new
version is committed to the repository. That way, a complete history of the
downloads is kept, but if the resource is not changing a lot, this history will
not take up an inordinate amount of space (because of the deduplication
properties of `git`).

If the resource gets borked, you can retrieve older versions using git. A function
is provided to list available pins (with or without history), but beyond that, the
user is expected to use `git` directly for more complex retrieval operations.

```{r, include=FALSE}
old_width <- options(width=130)
```

```{r}
list_gitpins()
list_gitpins(history = TRUE)
```

```{r, include=FALSE}
options(old_width)
```

## Installation

You can install the development version of `gitpins` like so:

``` r
pak::pak("torfason/gitpins")
# or alternatively
remotes::install_github("torfason/gitpins")
```

Note that `gitpins` uses the native pipe operator (`|>`) and so depends on `R
(>= 4.1.0)`. If this is an issue for you, holler and can probably be convinced
of changing it to make it compatible with older versions.

## Related packages and feedback

This package was inspired by the `pins` package, and in particular the `pin()` function.
However, that function stores the actual local file in a system location rather than
inside the project, so using it was not totally reliable. Also, it did not have 
the desired versioning properties. Thus, `gitpins` was born.

For feature requests, bugs, or other feedback, feel free to file an issue.


