---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = TRUE
)
library(gitpins)
stopifnot(!file.exists(here::here("gitpins")))
```

# gitpins

<!-- badges: start -->
[![R-CMD-check](https://github.com/torfason/gitpins/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/torfason/gitpins/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`r desc::desc_get("Description")`

## The Problem

You want to quickly and easily process an online resource using R functions,
some of which only accept local files. Thus you would like the following
properties for your workflow:

* Download to a local file
* But avoid downloading on every single run
* Refresh your data regularly from the online source
* Use a local copy if the online resource is not accessible
* Have the local copy be easily accessible in a predictable location
* Not ruin your local copy if the online version should change in a "bad" way

## The Solution

The `gitpins` package downloads a URL to a local file in the `gitpins` folder
(defaults to `here::here("gitpins")`, but can be configured using
`gp_options()`), and then returns the full file name name of the local file,
which can be passed as an argument to any function that expects to read such a
file.

## Installation

Install `gitpins` using `pak`:

```r
pak::pak("torfason/gitpins")
```

## Usage

```{r}
# Downloads on first try
pin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv") |> 
  read.csv() |> head()
```

You can maintain as many resources as you need:

```{r}
# Another resource
pin("https://vincentarelbundock.github.io/Rdatasets/csv/datasets/sunspot.month.csv") |> 
  read.csv() |> head()
```

The file is downloaded the first time you run `pin()` on a given URL (the
actual download is done with `curl::curl_download()`). After that, it checks to
see the age of the local file and re-downloads if it is to old. The default
refresh interval is 12 hours, but is configurable with a parameter.

Note that the return value of the `pin()` function is simply the full path to
the local copy of the file. You can therefore use `pin()` with the original
URL wherever you would have used the local path of the resource. The exact name
of the file is constructed in a deterministic way based on the URL
(specifically, the base name is the `digest()` of the URL).

```{r}
# Uses a cached copy if a recent one is available (start of the url changed for privacy)
pin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv") |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")
```

The refresh interval is configured with the `refresh_hours` parameter. Use
`refresh_hours=0` to force a download on every call, and `refresh_hours=Inf` to
always use the local copy (after the first download). A helper function,
`gp_dropper()` is provided for the case where a new version of the resource
"drops" at the same time every day. The function allows you to set a lower
interval in a given time window after the expected drop time, to maximize the
probability that an updated version gets downloaded quickly.

```{r}
# Force a reload by specifying zero refresh time
pin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv",
       refresh_hours = 0) |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")

# Always use local copy by specifying Inf refresh time
pin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv",
       refresh_hours = Inf) |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")

# Set a lower interval for a given time window after a resource update "drops"
pin("https://vincentarelbundock.github.io/Rdatasets/csv/openintro/country_iso.csv",
       refresh_hours = gp_dropper(drop_hour = 12, drop_tz = "US/Eastern")) |>
  gsub(pattern=".*/(gitpins/.*)", replacement="/home/user/project/\\1")
```

The `gitpins` directory is actually a local `git` repository, and each new
version is committed to the repository. That way, a complete history of the
downloads is kept, but if the resource is not changing a lot, this history will
not take up an inordinate amount of space (because of the deduplication
properties of `git`).

If the resource gets borked, you can retrieve older versions using git. A
function is provided to list available pins (with or without history), but
beyond that, the user is expected to use `git` directly for more complex
retrieval operations.

```{r}
withr::with_options(list(width = 130), {
  gp_list()
  gp_list(history = TRUE)
})
```


## Function Name Conflicts

For use with with another package that also defines a `pin()` function (such as
the `pins` package), the `conflicted` package comes highly recommended, but the
`exclude` option of the `library()` function is also a valid approach. In either
case, the `gp_pin()` function is provided as an alias for `pin()` so you don't
need to specify the full package name on each call:

### Using `conflicted`

```r
library(conflicted)
conflicts_prefer(pins::pin())
library(pins)
library(gitpins)
gp_pin(URL)
```

### Using `exclude`

```r
library(pins)
library(gitpins, exclude="pin")
gp_pin(URL)
```

## Related Packages, System Requirements, and Feedback

This package was inspired by the `pins` package, and in particular the
`pins::pin()` function. However, that function stores the actual local file in a
system location rather than inside the project, so using it did not prove
reliable. Furthermore, it did not have the desired versioning properties, and
finally, it is now defined as a legacy function and is not part of the new api
for that package. As a result, `gitpins` was born.

Note that `gitpins` uses the native pipe operator (`|>`) and so depends on `R
(>= 4.1.0)`. 

For feature requests, bugs, or other feedback, feel free to file an issue.


